---
title: "Strain_ABM"
author: "Rebecca"
date: "3 maj 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Quick notes about organisation 

In the first code chunck i load the packages necesary for the simulation to run.
I also use it for testing out the simulation. 

To actually run the simulation, you have to go down and define the function in the other R chunks. the Simulation() function comes first, but you should define that one last, as it is dependent on the other functions. 

```{r}
# Loading packages 
pacman::p_load(pacman, tidyverse)

# testing Simulation 
twodf <- Simulation(n_turns = 100,
                    n_agent = 5,0.75,
                    scrum = TRUE,
                    specs = "changing",
                    threshold = 2,
                    len = 10,
                    penalty = 0.1,
                    no = 0,
                    pos = 1,
                    neg = 0)
twodf[[1]]
twodf[[2]]

```

## Simulation() function

The Simulation() function gathers all the Work(), Sprint() and Specs() functions to comute the agent based model. 

- n_turns = the number of turns simulated 
- n_agent = the number of agents
- start_prob = the initial sucessrate (unmodulated by mental_health) of all the agents
- scrum = Scrum condition or not Scrum condition
- specs = specifying if the scope is changing or static
- threshold = the max distance agent_specs can drift from model_specs
- len = the length of the sprint in turns
- Penalty = the size of the penalty for project breakdown - a percentege of reduction in reward for each agent
- no = the probability of no change in scope 
- pos = the probability of a positive change in scope
- neg = the probability of a negative change in scope

```{r}
Simulation <- function(n_turns, n_agent, start_prob, scrum = TRUE, specs = "changing", threshold, len, penalty, no, pos, neg){
  
  # Making an array containing information about all agents
  agents <- array(0,c(n_turns,4,n_agent))
  dimnames(agents)[[2]] <- c("sucess", "effort", "reward", "mental_health")
  # Making a dataframe with additional variables
  values <- tibble("project_specs" = rep.int(0,n_turns), 
                   "agent_specs" = rep.int(0,n_turns), 
                   "sprint_turn" = rep_len(1:len,n_turns), 
                   "break" = rep.int(0,n_turns))
  
  #---> For the first round 
  # This happes in the scrum condition
  if(scrum == TRUE){
    # first turn in a sprint does not give any rewards
    agents[1,2,] <- agents[1,2,] + 1
  } else { # This happens in th non-scrum condition 
      for(i in 1:n_agent){
        # Defining the probability of a sucessful collaboration for the first turn
        s_prob <- start_prob
        # Determining if the collaboration is sucessful
        colab <- sample(c("sucess", "fail"), 1, prob = c(s_prob,(1-s_prob)))
        # Simulating collaboration
        if(colab == "sucess"){
          agents[1,1,i] <- 1
          agents[1,2,i] <- agents[1,2,i] + 1
          agents[1,3,i] <- agents[1,3,i] + 1
          agents[1,4,i] <- agents[1,3,i]/agents[1,2,i]
        } else {
          agents[1,2,i] <- agents[1,2,i] + 1 
          agents[1,4,i] <- agents[1,3,i]/agents[1,2,i]
        }
        
      }
  }

  # This happens in the changing specs condition 
  if(specs == "changing"){
    # Deciding if model_specs should be changed
    spec_change <- sample(c("none", "positive", "negative"), 1, prob = c(no, pos, neg))
    # if spec_chance is positive, 1 is added to model_specs
    if(spec_change == "positive"){
      values[1,1] <- 1
    }
    # if spec_chance is negative, 1 is subtracted from model_specs
    if(spec_change == "negative"){
      values[1,1] <- (- 1)
    }
  }
  
  #----> For loop for repeating repeating the same procedure for the rest of the turns
  for(i in 2:n_turns){
    # work is siulated and agent_frame is updated
    agents <- Work(start_prob, agents, i)
    # Sprints are simulated and values are udpated
    if(scrum == TRUE){
      values <- Sprint(len, values, i)
    }
    
    # Spacs are changed and both agent_frame and values are updated
    if(specs == "changing"){
      both <- Specs(threshold, agents, values, i, penalty, no, pos, neg)
      values <- both[[1]]
      agents <- both[[2]]
    }
  }
  
  # Retun an array with information about participants 
  return(list(values, agents))
}

```



## Work() function 

The Work() function simulates the agents working on the project. It is used in all conditions. 

- colab_prob = the defaoult probablility of a sucessful collaboration (ie. not modefied by mental health yet)
- array = array containing information about agents
- turn = current turn in the model (which row in the array should the function)

```{r}
Work <- function(colab_prob, arrai, turn) {
  
  # For loop repeating for each participant
  for(i in 1:dim(arrai)[3]){
    
    # Defining the probability of a sucessful collaboration
    if(arrai[turn-1,4,i] != 0){
      s_prob <- colab_prob*arrai[turn-1,4,i]
    } else {
      s_prob <- colab_prob
    }
    
    # Determining if the collaboration is sucessful
    colab <- sample(c("sucess", "fail"), 1, prob = c(s_prob,(1-s_prob)))
    
    # Simulating collaboration
    if(colab == "sucess"){
      arrai[turn,1,i] <- 1
      arrai[turn,2,i] <- arrai[turn-1,2,i] + 1
      arrai[turn,3,i] <- arrai[turn-1,3,i] + 1
      arrai[turn,4,i] <- arrai[turn,3,i]/arrai[turn,2,i]
    } else {
      arrai[turn,2,i] <- arrai[turn-1,2,i] + 1 
      arrai[turn,3,i] <- arrai[turn-1,3,i]
      arrai[turn,4,i] <- arrai[turn,3,i]/arrai[turn,2,i]
    }
  }
  
  # Retuning the modifyed dataframe
  return(arrai)
}

```


## Sprint() Function

Sprint() keeps track of the sprints and changes the model specifications in the Complex condition. It is only used in the Scrum conditions. It should happen at the end of a loop. 

- len = the length of the sprint in turns
- df = dataframe with information not connected to a specific agent
- turn = current turn in the model (which row in the array should the function)

```{r}
Sprint <- function(len, df, turn){
  
  # This happens at the last turn of the sprint
  if(df[turn, 3] == len){
    # agent_specs moves closer to project_specs if agent_specs is greater  
    if(df[turn-1, 2] > df[turn-1, 1]){
      df[turn, 2] <- df[turn-1, 2] - 1
    }
    # agent_specs moves closer to project_specs if agent_specs is smaller
    if(df[turn-1, 2] < df[turn-1, 2]){
      df[turn-1, 2] <- df[turn-1, 2] + 1
    }
  }
  
  # Function returns changed agent_specs and sprint_turn
  return(df)
}

```

## Specs() function

Specs() checks for drift in specifications and imposes a penalty on the simulation if the drift is over a certain threshold. It also chnages the specifications.

- threshold = the max distance agent_specs can drift from model_specs
- array = array containing information about agents
- df = dataframe with information not connected to a specific agent
- turn = current turn in the model (which row in the array should the function)
- no = the probability of no change in scope 
- pos = the probability of a positive change in scope
- neg = the probability of a negative change in scope
- Penalty = the size of the penalty for project breakdown - a percentege of reduction in reward for each agent

```{r}
Specs <- function(threshold, arrai, df, turn, penalty, no, pos, neg){
  # Checking if agent_specs has drifted
  spec_drift <- abs(df[turn-1,1] - df[turn-1,2])
  # This happens if agent_specs has drifted
  if(spec_drift >= threshold){
    # Penalty
    df[turn, 4] <- 1
    # ensuring that no reward score goes below 0 and calulating mental health with penalty
    for(i in 1:dim(arrai)[3]){
      arrai[turn, 3,] <- arrai[turn-1, 3,i] - round((arrai[turn-1, 3,i]*penalty))
      if(arrai[turn, 3, i] <= 0){
        arrai[turn, 3, i] <- 0
      }
      arrai[turn, 4, i] <- arrai[turn, 3, i]/arrai[turn, 2, i]
    }
    # Agent_specs are adapted to model_specs  
    df[turn, 2] <- df[turn-1, 1]
  }
  
  # Deciding if model_specs should be changed
  spec_change <- sample(c("none", "positive", "negative"), 1, prob = c(no, pos, neg))
  # if spec_chance is positive, 1 is added to model_specs
  if(spec_change == "positive"){
    df[turn, 1] <- df[turn-1, 1] + 1
  }
  # if spec_chance is negative, 1 is subtracted from model_specs
  if(spec_change == "negative"){
    df[turn, 1] <- df[turn-1, 1] - 1
  }
  return(list(df, arrai))
}

```


## Issues (notes for myself):

Right now, non-scrum seems to be doing much better than scrum (perhaps it is because the one round in the beginning without reward is not well proportioned with the consequence it creates)

I cannot visibly see when there is a change in project_specs - this might be a sign that I should change the probability of a change in specs. 

Right now the penalty for project breakdown is calculated individually, but maybe it shouldn't be? 

I want to ask someone about the work-flow of creating ABM's (ie. is it ok to change your model based on the results? How much?)

Still need to find out more about the implementation of the effort - reward imbalance model.

The mental_health statistic is hard to interpret as an outcome statistic. 

Also still need to rethink the entire structure of the model (I hate myself)

ALso rethink why the specs variable needs to go both ways and not just one. 

- (Solve today if you have more time) There are a bunch more variables I would like to be able to adjust without re-defining the function.

Individual sucess rates anyone? 

- (Solve today if you have time and brain-power for doing anything non-robotic) Should sprint or specs come first?

- (Solve if you have more time) plots. You should try to make some more plots. More plots is always good. 

## Extra code for de-bugging

```{r}
 # Making an array containing information about all agents
  agg <- array(0,c(10,4,3))
  dimnames(agg)[[2]] <- c("sucess", "effort", "reward", "mental_health")
  # Making a dataframe with additional variables
  val <- tibble("project_specs" = rep.int(0,10), 
                "agent_specs" = rep.int(0,10), 
                "sprint_turn" = rep_len(1:10,10), 
                "break" = rep.int(0,10))

obj <- Specs(2, agg, val, 2, 0.1, 1, 0, 0)
obj[[1]]
```


## Including Plots

If I want to include nice plots withing this markdown onece I have some nice resuts to plot

```{r pressure, echo=FALSE}
plot(pressure)
```


